<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{pxweb}
-->

PX-WEB API Interface for R
===========

This R package provides tools to access [PX-WEB
API](http://www.scb.se/Grupp/OmSCB/API/API-description.pdf). Your
[contributions](http://ropengov.github.com/contact.html) and [bug
reports and other feedback](https://github.com/ropengov/pxweb) are
welcome!

## Table of contents

[Installation](#installation) (Installation)  
[Examples](#examples) (Examples)  

## Available data sources and tools

[A number of organizations](http://www.scb.se/sv_/PC-Axis/Programs/PX-Web/PX-Web-examples/) use to distribute hierarchical data. You can browse the available data sets at:

* [Statistics Sweden](http://www.statistikdatabasen.scb.se/pxweb/en/ssd/) with [API Description](http://www.scb.se/Grupp/OmSCB/API/API-description.pdf)
* [Statistics Finland](http://tilastokeskus.fi/til/aihealuejako.html) [StatFi API Description](http://pxnet2.stat.fi/api1.html)
* [Other organizations using PX-WEB](http://www.scb.se/sv_/PC-Axis/Programs/PX-Web/PX-Web-examples/)

## <a name="installation"></a>Installation


Install the stable release version in R:

```{r install1, eval=FALSE}
install.packages("pxweb")
```

Install the latest version from github:

```{r install2, eval=FALSE}
install.packages("devtools")
devtools::install_github("pxweb","rOpenGov")
```

Test the installation by loading the library:

```{r test, message=FALSE, warning=FALSE, eval=TRUE}
library(pxweb)
```

A tutorial is included with the package with:
```r
vignette(topic="pxweb")
```

## <a name="examples"></a>Examples

Some examples on using the R tools to fetch px-web API data.

### Listing available pxweb apis

```{r apiparameters, message=FALSE, eval=TRUE}
library(pxweb)
print(api_catalogue()[1:2])
```

### Fetching data from PX-WEB API:

Interactive API query (not run):

```{r standardquery, message=FALSE, eval=FALSE}
# Get data from SCB (Statistics Sweden)
d <- interactive_pxweb(api = "api.scb.se")

# Fetching data from the swedish SCB (Statistics Sweden) pxweb API using the alias 'scb'
d <- interactive_pxweb(api = "scb", version = "v1", lang = "sv")

# Fetching data from statfi (Statistics Finland) using the alias 'statfi'
d <- interactive_pxweb(api = "statfi")
```

### Create a new PX-WEB API:

If a pxweb api is missing in the api catalogue, you can just add the api yourself in the following way.

```{r add_api, message=FALSE, eval=FALSE}
# Create a pxweb api object
my_api <- 
  pxweb_api$new(api = "foo.bar", # PXWEB api name
                alias = "foo", # Alias for the api, can be ignored
                url = "http://api.foo.bar/[version]/[lang]",
                description = "My own pxweb api",
                languages = "en", # Languages
                versions = "v1", # Versions         
                calls_per_period = 1,
                period_in_seconds = 2, 
                max_values_to_download = 10)

# Test that the api works (in this case it does not)
my_api$test_api()

# Add the api to the api catalogue
my_api$write_to_catalogue()

```


## Licensing and Citations

This work can be freely used, modified and distributed under the open license specified in the [DESCRIPTION file](https://github.com/rOpenGov/pxweb/blob/master/DESCRIPTION).

Kindly cite the work as follows

```{r citation, message=FALSE, eval=TRUE}
citation("pxweb")
```

## About the API

The data in this RESTful API consists of a metadata part and a data
part. Metadata is structured in a hierarchical node tree, where each
node contains information about subnodes that are below it in the tree
or, if the nodes is at the bottom of the tree structure, the data
referenced by the node as well as what dimensions are available for
the data at that subnode.


## Session info

This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```




