<!--
%\VignetteEngine{knitr::knitr}
%\VignetteIndexEntry{pxweb Markdown Vignette made with knitr}
-->

PX-WEB API Interface for R
===========

This R package provides tools to access [PX-WEB
API](http://www.scb.se/Grupp/OmSCB/API/API-description.pdf). Your
[contributions](http://ropengov.github.com/contact.html) and [bug
reports and other feedback](https://github.com/ropengov/pxweb) are
welcome!


## Available data sources and tools

[Installation](#installation) (Installation)  
[Examples](#examples) (Examples)  

[A number of organizations](http://www.scb.se/sv_/PC-Axis/Programs/PX-Web/PX-Web-examples/) use to distribute hierarchical data. You can browse the available data sets at:

* [SCB](Source: [SCB](http://www.statistikdatabasen.scb.se/pxweb/en/ssd/) (Statistiska centralbyrån) (Sweden)
* [Statistics Finland](http://tilastokeskus.fi/til/aihealuejako.html) (Statistics Finland)
* [Other organizations with PX-WEB API](http://www.scb.se/sv_/PC-Axis/Programs/PX-Web/PX-Web-examples/)

## <a name="installation"></a>Installation

Install the stable release version in R:

```{r install, eval=FALSE}
install.packages("pxweb")
```

Test the installation by loading the library:

```{r test, message=FALSE, warning=FALSE, eval=TRUE}
library(pxweb)
```

We also recommend setting the UTF-8 encoding:

```{r locale, eval=FALSE}
Sys.setlocale(locale="UTF-8") 
```

## <a name="examples"></a>Examples

Some examples on using the R tools to fetch px-web API data.

### Listing available database parameters

```{r apiparameters, message=FALSE, eval=TRUE}
library(pxweb)
print(api_parameters())
```

### Fetching data from [Statistics Finland](http://www.stat.fi/org/avoindata/api.html) PX-WEB API:

Interactive API query (not run):

```{r standardquery, message=FALSE, eval=FALSE}
baseURL <- base_url("statfi", "v1", "fi")
d <- findData(baseURL)
```


## Licensing and Citations

This work can be freely used, modified and distributed under the open license specified in the [DESCRIPTION file](https://github.com/rOpenGov/pxweb/blob/master/DESCRIPTION).

Kindly cite the work as follows

```{r citation, message=FALSE, eval=TRUE}
citation("pxweb")
```

## Session info

This vignette was created with

```{r sessioninfo, message=FALSE, warning=FALSE}
sessionInfo()
```




